### 2020-01-28 ChapterImplementation/chapterImplementation.tex
Fixed: Language in Implementation

\chapter{Implementation}
In this section will discuss our implementation of AP, The following will be looked at in detail: \begin{itemize}
	\item An overview of the system architecture: How the servers are deployed in the cloud and how they communicate with a client; a breakdown of the components of each server; and potential deployments to be investigated in the future.
	\item Aura Implementation
	\item The algorithms used by the servers to implement AP.
	\item The life cycle of a simulated object as a finite state machine (FSM).
	\item The features and implementation details of the visualiser, including a description of its features and future work to be carried out on the visualiser.
\end{itemize}

@@

Any topology more complex than columns will result in corners being present in the topology. The simplest corner that can exist is between three servers, as shown in Fig. \ref{3ServerCornerLayout}. All three server must communicate when a corner between three servers exists, this is required for the solution to the corner case (as described in \ref{CornerCase}). For example, in  Fig. \ref{3ServerCornerLayout}, a cluster of objects being hosted by server 2 may exist across the boundary between server 0, objects on the edge of the cluster may be positioned deep into server 0's region. In addition a cluster of objects being hosted on server 1 may exist across the boundary between server 1 and server 0 with some objects in the cluster being deep into server 0's region. Without the auras from the server 0 - server 1 boundary being sent to server 2 and vice versa, the two clusters will be unaware of each other's existence. From Server 0's perspective, the two clusters may have overlapping auras, and so should be being simulated on the same server. In order to prevent this, the auras from the boundaries of all neighbours need to be exchanged.

@@

Overseer nodes could be located at the edge, geographically closer to the clients, as illustrated in \ref{EdgeOverseer}. The advantage of this approach is, there is a reduced latency between clients and overseers, meaning lower delay times between clients and the workload being carried out by the overseers, such as game logic. The disadvantage of this approach is there is higher latency between overseer nodes, which are located far apart, geographically.

@@

\section{Aura Implementation}
How the auras in AP are implemented will now be described at a high-level, before a formal description of the algorithms using pseudo-code is given in \ref{Algorithms}.
Auras are implemented through the use of trigger volumes (non-physically interacting volumes, that generate trigger collision callbacks). Two trigger volumes are used for different purposes. 
The first is used for auras that are being received from remote servers. When an aura is received, a trigger volume, consisting of a sphere, is created at the corresponding position with a given radius. The second trigger volume used is by the server-region boundaries. Each server-region boundary has a trigger volume that extends inside the server's region, it extends the same distance as the aura radius, minus the bounding sphere of an object. This is used to determine when an object should begin and stop having its aura sent to neighbouring servers and avoid the need for per-object checks for proximity to the server-region boundary. When an object collides with the aura trigger boundary, the aura begins being sent, when an object leaves the aura trigger boundary, an aura delete message is sent.
Each server-region boundary has a trigger volume, objects may collide with multiple aura trigger boundaries, meaning auras can be sent to multiple servers, as illustrated in Fig. \ref{AuraCorner}.

\begin{figure}
	\centering
	\scalebox{0.75}{\input{Figs/AuraCorner.pdf_tex}}
	\caption{Aura Projection at a 3-way corner intersection. When an object collides with a server aura trigger boundary, the aura begins being sent to the corresponding neighbouring server. In this scenario, an object from server 0 is colliding with two aura trigger boundaries from servers 1 and 2, meaning the aura is sent to both servers 1 and 2. $x$ is the aura distance beyond the boundary sphere of an object and the distance used by the aura trigger boundaries.}
	\label{AuraCorner}
\end{figure}
%TODO: Boundary has a trigger volume, illustrate with diagram. Is this mentioned elsewhere?

@@

When a server begins communicating an aura, a message is added to the aura creation/update buffer. This message consists of the position of the centre of the aura, the radius of the aura. % and a flag signifying whether or not this is a creation message or not. Flag is not currently used as often causes issue in the corner case, possibly due to multiple types of messages concerncing one object being sent and recieved in the same tick.
The contents of the buffer are then broadcast to all neighbouring servers. When aura creation/update messages are received and an aura for that object does not already exist, a new aura, a sphere trigger volume, is created at that position with the given radius.
Whenever an object projecting an aura moves, and update message is added to the aura creation/update buffer, containing the new position and radius of the aura (aura radius is currently fixed, but may become dynamic in the future). The contents of the buffer are broadcast to all neighbouring servers. When the message is received, the positions of the auras are updated. 
If an object leaves aura trigger boundary, either by moving away from the boundary or by being migrated to another server, a delete aura message is added to the aura delete buffer. The contents of the buffer are then broadcast to all neighbouring servers. When the message received, the aura trigger volume is removed from the simulation.

When objects collide with an aura trigger volume from an object on a remote server, a callback from PhysX is generated, when processed by AP, the colliding object is added to the migration buffer for the server the aura is being sent from. In addition, a recursive search is carried out to identify all objects within the cluster (a group of objects with overlapping auras). The entire cluster is added to the migration buffer, so that all objects in the cluster are migrated at once. The contents of the buffer are then sent to the corresponding server.

@@

$BT$ is called when an object traverses a boundary. In order to prevent `islands' forming (for example Objects H, I and J in Fig. \ref{AuraProj}), a recursive search is carried out to determine is an object is part of an island or not. If an object is found to not be part of an island, the entire cluster of objects is added to the send buffer, otherwise no action is taken.

@@

We will now discuss the state of an object in the context of AP, i.e. not position and rotation of an object, but the state of object as it is being treated by AP.

@@

From the Intersecting Boundary state, an object remains in this state until on of the following occurs: the object is no longer intersecting the server boundary as is contained entirely within the host region (triggered through a PhysX callback, the OBE algorithm is run), in which case the object moves to the Host Region state; the object is no longer intersecting the server boundary and has fully traversed the boundary, in which case the object moves to the Boundary Traversed state; or the object collides with a remote object's aura (triggered through a PhysX callback, the AC algorithm is run), in which case it is immediately moved into the Sent State.

@@

\section{The Visualiser}
%We discuss the features and implementation of the visualiser, including server and client created replicas, static objects and interactivity. The challenges of multiple servers are explained and our solutions justified. Screenshots from demo scenes have been included.
We will now discuss the details of the visualiser, both the features and implementation details.

@@

\subsection{Terminology}
Before discussing specific details of the visualiser, we must first define terms used.
\begin{itemize}
	\item The visualiser: The software application that is capable of connecting to AP servers and display replicas of objects on the servers.
	\item The editor: This refers to the Unreal Engine Editor. The version of the editor used for the visualiser is UE4.19.2
	\item Edit-time: When scenes are being edited in the Unreal Engine Editor, before the visualiser has been packaged.
	\item Run-time: When the visualiser is running
	\item Player: The in-visualiser representation of the user. Has a position and rotation. The camera for rendering is attached to the player
	\item Replica object: A representation displayed in the visualiser of an object being simulated on a server, which is intended to replicate the physical state of the object on the server.
\end{itemize}

@@

Replica objects can be created in one of two ways. Either by the server or by the client. We will now discuss the implementation details of each type of replica creation.

@@

\subsection{Client-created static objects}\label{Client-CreatedReplicas}
In addition to replica objects, static object can also be created by the client. As these objects are static, there is no need for their states to be replicated as they do not change over the course of the simulation. Static object work simply by the client sending the geometry, position and rotation details of the static object to all servers and the servers then re-create the static object. These static objects then remain unchanged throughout the simulation, requiring no updates to the state to be communicated to the client. Static objects are created on all servers, this could be optimised so only static objects within a server's region are created, however, AP can lead to clusters of objects being simulated outside of their host server's region, which may interact with static objects that do not exist on their host server. In addition, the cost of simulating static objects tends to be very low, as they do not require any updates as they are not dynamic. In the future, region boundaries may change; a server may become responsible for a static object previously residing in another server's region, which would then have to be communicated when the region boundaries changed, by all servers simulating all static objects, this avoids the need for extra communication when server boundaries change. With these trade-offs in mind, we have elected to create all static objects on all servers, when static objects are used in a simulation.

@@

The visualiser can create rigid-bodies, including rigid-bodies with multiple shapes, however, joints (constraints) are not yet supported. Joints have the added complexity of needing references to multiple objects. In order for these to be replicated on the server, these references would need to be kept and correctly refer to the corresponding objects after being transferred to the servers. This could be achieved through the joint replica containing unique references to the object replicas, which are identical on both client and server.

>>>>

\chapter{Implementation}
The implementation of AP will be discussed in this section. The following points will be looked at in detail: 
\begin{itemize}
	\item An overview of the system architecture: How the servers are deployed in the cloud and how they communicate with a client; a breakdown of the components of each server; and potential deployments to be investigated in the future.
	\item Aura Implementation
	\item The algorithms used by the servers to implement AP.
	\item The life cycle of a simulated object as a finite state machine (FSM).
	\item The features and implementation details of the visualiser, including a description of its features and future work to be carried out on the visualiser.
\end{itemize}

@@

Any topology more complex than columns will result in corners being present in the topology. The simplest corner that can exist is between three servers, as shown in Fig. \ref{3ServerCornerLayout}. All three servers must communicate when a corner between three servers exists, this is required for the solution to the corner case (as described in \ref{CornerCase}). For example, in  Fig. \ref{3ServerCornerLayout}, a cluster of objects being hosted by server 2 may exist across the boundary between server 0, objects on the edge of the cluster may be positioned deep into server 0's region. In addition a cluster of objects being hosted on server 1 may exist across the boundary between server 1 and server 0 with some objects in the cluster being deep into server 0's region. Without the auras from the server 0 - server 1 boundary being sent to server 2 and vice versa, the two clusters will be unaware of each other's existence. From Server 0's perspective, the two clusters may have overlapping auras, and so should be being simulated on the same server. In order to prevent this, the auras from the boundaries of all neighbours need to be exchanged.

@@

Overseer nodes could be located at the edge, geographically closer to the clients, as illustrated in \ref{EdgeOverseer}. The advantage of this approach is, there is a reduced latency between clients and overseers, meaning lower delay times between clients and the workload being carried out by the overseers, such as game logic. The disadvantage of this approach is that there is higher latency between overseer nodes, which are located far apart, geographically.

@@

\section{Aura Implementation}
How the auras in AP are implemented will now be described at a high-level, before a formal description of the algorithms using pseudo-code is given in \ref{Algorithms}.
Auras are implemented through the use of trigger volumes (non-physically interacting volumes, that generate trigger collision callbacks). Two trigger volumes are used for different purposes. 
The first is used for auras that are being received from remote servers. When an aura is received, a trigger volume, consisting of a sphere, is created at the corresponding position with a given radius. The second trigger volume used is by the server-region boundaries. Each server-region boundary has a trigger volume that extends inside the server's region, it extends the same distance as the aura radius, minus the bounding sphere of an object. This is used to determine when an object should begin and stop having its aura sent to neighbouring servers and avoid the need for per-object checks for proximity to the server-region boundary. When an object collides with the aura trigger boundary, the aura begins to be sent, when an object leaves the aura trigger boundary, an aura delete message is sent.
Each server-region boundary has a trigger volume, objects may collide with multiple aura trigger boundaries, meaning auras can be sent to multiple servers, as illustrated in Fig. \ref{AuraCorner}.

\begin{figure}
	\centering
	\scalebox{0.75}{\input{Figs/AuraCorner.pdf_tex}}
	\caption{Aura Projection at a 3-way corner intersection. When an object collides with a server aura trigger boundary, the aura begins to be sent to the corresponding neighbouring server. In this scenario, an object from server 0 is colliding with two aura trigger boundaries from servers 1 and 2, meaning the aura is sent to both servers 1 and 2. $x$ is the aura distance beyond the boundary sphere of an object and the distance used by the aura trigger boundaries.}
	\label{AuraCorner}
\end{figure}
%TODO: Boundary has a trigger volume, illustrate with diagram. Is this mentioned elsewhere?

@@

When a server begins communicating an aura, a message is added to the aura creation/update buffer. This message consists of the position of the centre of the aura, the radius of the aura. % and a flag signifying whether or not this is a creation message or not. Flag is not currently used as often causes issue in the corner case, possibly due to multiple types of messages concerncing one object being sent and recieved in the same tick.
The contents of the buffer are then broadcast to all neighbouring servers. When aura creation/update messages are received and an aura for that object does not already exist, a new aura, a sphere trigger volume, is created at that position with the given radius.
Whenever an object projecting an aura moves, and update message is added to the aura creation/update buffer, containing the new position and radius of the aura (aura radius is currently fixed, but may become dynamic in the future). The contents of the buffer are broadcast to all neighbouring servers. When the message is received, the positions of the auras are updated. 
If an object leaves aura trigger boundary, either by moving away from the boundary or by being migrated to another server, a delete aura message is added to the aura delete buffer. The contents of the buffer are then broadcast to all neighbouring servers. When the message is received, the aura trigger volume is removed from the simulation.

When objects collide with an aura trigger volume from an object on a remote server, a callback from PhysX is generated. When these callbacks are processed by AP, the colliding object is added to the migration buffer for the server the aura is being sent from. In addition, a recursive search is carried out to identify all objects within the cluster (a group of objects with overlapping auras). The entire cluster is added to the migration buffer, so that all objects in the cluster are migrated at once. The contents of the buffer are then sent to the corresponding server.

@@

$BT$ is called when an object traverses a boundary. In order to prevent `islands' forming (for example Objects H, I and J in Fig. \ref{AuraProj}), a recursive search is carried out to determine if an object is part of an island or not. If an object is found to not be part of an island, the entire cluster of objects is added to the send buffer, otherwise no action is taken.

@@

The state of an object in the context of AP will now be discussed, i.e. not the physical state of an object (position and rotation), but the state of object as it is being treated by AP.

@@

From the Intersecting Boundary state, an object remains in this state until one of the following occurs: the object is no longer intersecting the server boundary as is contained entirely within the host region (triggered through a PhysX callback, the OBE algorithm is run), in which case the object moves to the Host Region state; the object is no longer intersecting the server boundary and has fully traversed the boundary, in which case the object moves to the Boundary Traversed state; or the object collides with a remote object's aura (triggered through a PhysX callback, the AC algorithm is run), in which case it is immediately moved into the Sent State.

@@

\section{The Visualiser}
%We discuss the features and implementation of the visualiser, including server and client created replicas, static objects and interactivity. The challenges of multiple servers are explained and our solutions justified. Screenshots from demo scenes have been included.
The details of the visualiser, both the features and implementation details, will now discussed.

@@

\subsection{Terminology}
Before discussing specific details of the visualiser, the terms used must be defined.
\begin{itemize}
	\item The visualiser: The software application that is capable of connecting to AP servers and display replicas of objects on the servers.
	\item The editor: This refers to the Unreal Engine Editor. The version of the editor used for the visualiser is UE4.19.2
	\item Edit-time: When scenes are being edited in the Unreal Engine Editor, before the visualiser has been packaged.
	\item Run-time: When the visualiser is running
	\item Player: The in-visualiser representation of the user. Has a position and rotation. The camera for rendering is attached to the player
	\item Replica object: A representation displayed in the visualiser of an object being simulated on a server, which is intended to replicate the physical state of the object on the server.
\end{itemize}

@@

Replica objects can be created in one of two ways. Either by the server or by the client. The implementation details of each type of replica creation will now be discussed.

@@

\subsection{Client-created static objects}\label{Client-CreatedReplicas}
In addition to replica objects, static objects can also be created by the client. As these objects are static, there is no need for their states to be replicated as they do not change over the course of the simulation. Static object work simply by the client sending the geometry, position and rotation details of the static object to all servers and the servers then re-create the static object. These static objects then remain unchanged throughout the simulation, requiring no updates to the state to be communicated to the client. Static objects are created on all servers, this could be optimised so only static objects within a server's region are created, however, AP can lead to clusters of objects being simulated outside of their host server's region, which may interact with static objects that do not exist on their host server. In addition, the cost of simulating static objects tends to be very low. They do not require any updates as they are not dynamic. In the future, region boundaries may change; a server may become responsible for a static object previously residing in another server's region, which would then have to be communicated when the region boundaries changed, by all servers simulating all static objects, this avoids the need for extra communication when server boundaries change. With these trade-offs in mind, we have elected to create all static objects on all servers, when static objects are used in a simulation.

@@

The visualiser can create rigid-bodies, including those with multiple shapes, however, joints (constraints) are not yet supported. Joints have the added complexity of needing references to multiple objects. In order for these to be replicated on the server, these references would need to be kept and correctly refer to the corresponding objects after being transferred to the servers. This could be achieved through the joint replica containing unique references to the object replicas, which are identical on both client and server.
